---
title: "Reproducible Research: Peer Assessment 1"
author: "Nikhil Prakash"
---
## Overview
This assignment makes use of data from a personal activity monitoring device.
This device collects data at 5 minute intervals throughout the day. The data 
consists of two months of data from an anonymous individual collected during 
the months of October and November, 2012 and include the number of steps taken 
in 5 minute intervals each day.

```{r results='hide'}
library(dplyr)
library(ggplot2)
```
## Loading and preprocessing the data
```{r}
actData <- read.csv("activity.csv", header = TRUE)
head(actData)
```

## What is mean total number of steps taken per day?
```{r}
steps <- aggregate(actData$steps, by = list(Date = actData$date), FUN = sum)
names(steps)[names(steps) == "x"] <- "Total"
temp <- as.Date(steps$Date, "%Y-%m-%d")
steps$Date <- format(temp, format = "%m-%d")
head(steps)
```
```{r}
ggplot(data = na.omit(steps), aes(Total)) + geom_histogram(binwidth = 1500, colour = "white") +
    xlab("Total Number of Steps Taken Each Day") + ylab("Count") +
    ggtitle("Histogram of the Total Number of Steps Taken Each Day")
```
```{r}
mean(na.omit(steps$Total))
median(na.omit(steps$Total))
```

## What is the average daily activity pattern?
```{r}
pattern <- aggregate(steps ~ interval, data = actData, FUN =mean)
ggplot(data = pattern, aes(x = interval, y = steps)) + geom_line() +
    xlab("Time Intervals (5 Minutes interval)") + ylab("Total Number of Steps") +
    ggtitle("Average Number of Steps Taken in 5-Minute Interval")
```

```{r}
head(pattern)
```
```{r}
pattern[which(pattern$steps == max(pattern$steps)),]
```

## Imputing missing values
```{r}
sapply(X = actData, FUN = function(x) sum(is.na(x)))
```
## Replace with Mean
```{r}
replacewithmean <- function(x) replace(x, is.na(x), mean(x, na.rm = TRUE))
meandata <- actData%>% group_by(interval) %>% mutate(steps= replacewithmean(steps))
head(meandata)
```
```{r}
newdata <- as.data.frame(meandata)
```
```{r}
head(newdata)
```
```{r}
summary(newdata)
```
```{r}
stepsByDay <- tapply(meandata$steps, meandata$date, sum)
qplot(stepsByDay, xlab='Total steps per day', ylab='Frequency', binwidth=500)
```
## Compare Mean and Median 
```{r}
mean(na.omit(steps$Total))
mean(na.omit(steps$Total))
median(na.omit(steps$Total))
median(na.omit(steps$Total))
```

## Are there differences in activity patterns between weekdays and weekends?

```{r}
meandata$date <- as.Date(meandata$date)
meandata$weekday <- weekdays(meandata$date)
meandata$weekend <- ifelse(meandata$weekday=="Saturday" | meandata$weekday=="Sunday", "Weekend", "Weekday" )
meandataweekendweekday <- aggregate(meandata$steps , by= list(meandata$weekend, meandata$interval), na.omit(mean))
names(meandataweekendweekday) <- c("weekend", "interval", "steps")
ggplot(meandataweekendweekday, aes(x=interval, y=steps, color=weekend)) + geom_line()+
facet_grid(weekend ~.) + xlab("Interval") + ylab("Mean of Steps") +
    ggtitle("Comparison of Average Number of Steps in Each Interval")
```

## Conclusion:
 From the two plots it seems that the test object is more active earlier in the day
 during weekdays compared to weekends, but more active throughout the weekends compared
 with weekdays.
